% Separate title page for double-blind submission to the European Journal for Philosophy of Science.
% This file is uploaded separately and is NOT part of the anonymised manuscript.
\documentclass[12pt,a4paper]{article}

\usepackage{hyperref}
\usepackage[margin=2.5cm]{geometry}

\begin{document}

  \begin{center}
    {\LARGE Non-Injectivity as a Structural Necessity of Genuine Emergence:\\[0.3em]
    A No-Go Theorem for Faithful Emergent Descriptions}\\[2em]

    {\large J\'er\^ome Beau}\\[0.5em]
    Independent Researcher, Paris, France\\
    \texttt{jerome.beau@cosmochrony.org}\\
    ORCID: \url{https://orcid.org/0009-0001-7697-7868}
  \end{center}

  \vspace{2em}

  \section*{Corresponding author}

  J\'er\^ome Beau, \texttt{jerome.beau@cosmochrony.org}.

  \section*{Acknowledgements}

  This work is part of a research programme on projective emergent physics.
  The author acknowledges the use of large language models as a supportive tool for refining language,
  structure, and internal consistency during the development of this manuscript.
  All conceptual contributions, theoretical choices, and interpretations remain the sole responsibility of the
  author.

  \section*{Statements and Declarations}

  \paragraph{Availability of data and material.}
  Data availability statement: not applicable.

  \paragraph{Funding.}
  No funding was received to assist with the preparation of this manuscript.

  \paragraph{Competing interests.}
  The author has no competing interests to declare that are relevant to the content of this article.

  \paragraph{Use of large language models.}
  Large language models were used as a supportive tool for refining language, structure, and internal
  consistency; all conceptual contributions, theoretical choices, and interpretations remain the sole
  responsibility of the author.

  \paragraph{Author contributions.}
  The sole author conceived the result, developed the proofs, and wrote the manuscript.

\end{document}
