\documentclass[12pt,a4paper]{article}

\usepackage{amsmath,amsthm,amssymb}
\usepackage[round,authoryear]{natbib}
\usepackage{hyperref}
\usepackage{doi}
\usepackage{microtype}
\usepackage[margin=2.5cm]{geometry}

% Anonymised variant for double-blind review (EJPS).
% Author information, acknowledgements and declarations are on the separate
% title page (EJPSTitlePage.tex), per the journal's double-blind procedure.
\hypersetup{pdfauthor={}, pdftitle={Non-Injectivity as a Structural Necessity of Genuine Emergence}}

\newtheorem{theorem}{Theorem}
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{proposition}[theorem]{Proposition}
\newtheorem{corollary}[theorem]{Corollary}
\newtheorem{example}[theorem]{Example}
\newtheorem{definition}{Definition}
\newtheorem{remark}{Remark}

\newcommand{\bstar}{\beta^*}

\title{Non-Injectivity as a Structural Necessity of Genuine Emergence:\\
A No-Go Theorem for Faithful Emergent Descriptions}

% Anonymised for double-blind review; see separate title page.
\author{}

\date{}

\begin{document}

  \maketitle

  \begin{abstract}
    We prove that any genuinely emergent (i.e.\ non-isomorphic effective) physical
    description is necessarily non-injective and intrinsically compressive.
    The argument is purely structural: given a surjective projection
    $\Pi : \Omega \to \mathcal{O}$ such that $\mathcal{O}$ carries no
    independent microstate-resolving structure beyond what $\Pi$ provides,
    and such that the effective level is not structurally isomorphic to the
    fundamental level, then $\Pi$ cannot be injective and the associated
    projection entropy $S_\Pi > 0$.
    The proof uses only surjectivity, informational completeness of $\Pi$ for
    observables, and the non-isomorphism condition; it is independent of any
    specific physical framework.
    The result constitutes a \emph{no-go theorem for faithful emergence}: a
    surjective projection that is both informationally faithful and genuinely
    emergent cannot exist.
    As a corollary, descriptive redundancy, non-injectivity of $\Pi$, and
    $S_\Pi > 0$ are co-extensive in any projective emergent framework, and any
    non-trivial gauge structure requires these properties.
    We provide a categorical reformulation (genuine emergence corresponds to
    a non-faithful functor) and an informational one (genuine emergence forces
    a strict, not merely non-negative, data-processing information loss).
    The result is illustrated by a concrete spin model and discussed in
    relation to the projective incompleteness of $\Lambda$CDM and the
    Standard Model, whose fundamental constants are identified as invariants
    of an unspecified projection $\Pi$ not accessible from the effective level
    alone.
    A further corollary establishes that non-injectivity is recursive: any
    tower of genuinely distinct emergent levels carries one independent
    non-injective projection per stratum, with entropy accumulating across
    levels --- implying, wherever colour-neutral hadrons arise as genuinely
    emergent from coloured degrees of freedom, the structural non-observability
    of free colour: the structural, not dynamical, content of confinement.
  \end{abstract}

  \paragraph{Keywords.}
    emergence, non-injectivity, projection, gauge symmetry, information loss,
    effective theories.

    \clearpage
    \tableofcontents

    \clearpage

  \section{Introduction}
  \label{sec:intro}

  Many physical frameworks assume the existence of two distinct levels of
  description: a fundamental level $\Omega$ of underlying configurations, and
  an effective level $\mathcal{O}$ of observable states.
  The relationship between these levels is typically encoded in a projection
  map $\Pi : \Omega \to \mathcal{O}$.

  A recurring feature of such frameworks is that $\Pi$ is many-to-one:
  distinct configurations $\omega \in \Omega$ may correspond to the same
  observable state $o \in \mathcal{O}$.
  This feature is closely related to coarse-graining and renormalisation
  procedures, where effective descriptions arise through systematic loss
  of microscopic information~\citep{Wilson1975,Jaynes1957,GellMannHartle1993}.
  More generally, the relation between emergence and information
  compression has been discussed in both physical and philosophical
  contexts, from Anderson's classic statement of the autonomy of emergent
  levels~\citep{Anderson1972} to recent systematic
  classifications~\citep{Butterfield2011,Ellis2012,CarrollParola2024}.
  This non-injectivity has been identified as the structural origin of gauge
  symmetry~\citep{Cosmochrony}, Bell-inequality violations~\citep{Beau2026b},
  and black hole entropy~\citep{Beau2026c}, and is broadly related to
  redundancy structures appearing in constrained and coarse-grained
  descriptions.
  In each case, however, non-injectivity has been introduced as a feature of
  a specific model rather than derived as a universal necessity.
  What is missing across these frameworks is a structural necessity result:
  while non-injective mappings appear operationally, no general theorem
  establishes that they are unavoidable for emergence itself.

  The same lacuna is visible from the philosophical side of the emergence debate.
  Discussions of emergence standardly distinguish weak emergence, on which emergent behaviour is in principle
  derivable from the base level but only by explicit simulation~\citep{Bedau1997}, from strong emergence, on
  which the emergent level carries facts not deducible from the base level even in
  principle~\citep{Chalmers2006}.
  Both notions are characterised epistemically --- by what can be derived, deduced, or simulated --- and the
  persistent difficulty of the debate has been to say what, in the \emph{structure} of an inter-level
  relation, distinguishes genuine emergence from mere redescription~\citep{Kim1999, Butterfield2011}.
  The present result addresses precisely this question in the restricted but exactly analysable setting of
  projective frameworks: when the inter-level relation is a surjective projection and the effective
  observables are entirely defined by it, genuine novelty (non-isomorphism of the two levels) is
  \emph{equivalent} to a structural property of the map itself --- non-injectivity, and hence strictly
  positive information loss.
  The criterion is neither epistemic nor dynamical: it does not mention derivability, simulation, or any
  observer, and it is decidable at the level of the map $\Pi$ alone.

  The present note supplies this missing result.
  We show that non-injectivity is not a modelling choice but a structural
  consequence of genuine emergence: any surjective map $\Pi$ whose observables
  are entirely defined by $\Pi$, and whose effective level is not isomorphic
  to the fundamental level, must be non-injective.
  The proof is elementary and framework-independent.
  A further corollary (Section~\ref{sec:towers}) shows that non-injectivity is
  recursive: the same structural necessity applies at every stratum of a tower
  of emergent descriptions, with entropy accumulating across levels.

  \subsection*{A preliminary observation: standard cosmology as an effective description}

    Before stating our results, we draw attention to a structural feature of
    Big Bang cosmology that motivates the theorem.

    In the standard cosmological model $\Lambda$CDM, an effective spacetime
    geometry $g_{\mu\nu}$ is well defined and descriptively valid for
    $t > t_\mathrm{Planck}$.
    For $t \lesssim t_\mathrm{Planck}$, the geometric description ceases to be
    valid: the metric diverges, the Einstein equations break down, and
    $g_{\mu\nu}$ is not a well-defined object.
    Yet the large-scale geometry that exists for $t \gg t_\mathrm{Planck}$ is
    stable, continuous, and physically meaningful.

    These three facts suggest that $g_{\mu\nu}$ should be regarded as an
    effective object.
    A description that fails in a given regime is naturally interpreted as
    arising from a more fundamental structure whose validity extends beyond
    that regime.
    From this perspective, the breakdown of $g_{\mu\nu}$ indicates the limit of
    applicability of the geometric description, rather than a failure of
    physical existence.

    If this line of interpretation is adopted, one may view $g_{\mu\nu}$ as
    arising from a more primitive structure $\Omega$, through a projection
    $\Pi$ that becomes well-defined when appropriate conditions for
    projectability are met.

    \begin{remark}[Heuristic status of this argument]
      \label{rem:lcdm-heuristic}
      The implication ``$g_{\mu\nu}$ breaks down $\Rightarrow$ $\Pi$ exists and
      is non-injective'' is a heuristic rather than a logical consequence of
      Theorem~\ref{thm:main}.
      Theorem~\ref{thm:main} applies to frameworks that explicitly posit $\Omega$
      and $\Pi$; it does not construct them from the mere existence of a
      singularity.
      The argument here is interpretive: it identifies the breakdown of
      $g_{\mu\nu}$ as a structural signal that the geometric description is
      effective, and suggests that any complete successor theory would need to
      supply the missing $\Omega$ and $\Pi$.
      This does not constitute a critique of the standard model, but a
      reformulation of its domain of validity in structural terms.
      Once such a successor is posited --- and once it satisfies
      Definitions~\ref{def:defined-by-Pi} and~\ref{def:non-trivial} ---
      Theorem~\ref{thm:main} applies and forces non-injectivity.
    \end{remark}

    In this sense, any theory that admits a regime where its own description
    ceases to be valid can be interpreted as an effective description with a
    restricted domain of applicability.
    Such a theory does not necessarily specify the underlying level explicitly,
    but its structure is compatible with an emergent interpretation.

    \begin{quote}
      \emph{Any cosmological framework admitting a regime where spacetime geometry
      ceases to be well-defined can be consistently interpreted as an effective
      description, suggesting the existence of a more fundamental level from which
      geometry arises.
      The absence of an explicit formulation of this underlying level does not
      eliminate this interpretation, but leaves it structurally incomplete.}
    \end{quote}

    From this viewpoint, the absence of a geometric description for
    $t \lesssim t_\mathrm{Planck}$ should be understood as delimiting the
    regime of validity of the spacetime description, rather than as implying
    the absence of an underlying structure.

    If a projection $\Pi$ relating such a fundamental level $\Omega$ to the
    effective observables $\mathcal{O}$ is introduced and satisfies the
    conditions of Section~\ref{sec:framework}, Theorem~\ref{thm:main} implies
    that it must be non-injective.
    The non-injectivity is therefore not a model-dependent feature, but a
    structural necessity of any complete projective-emergent framework.

    \textbf{Key message.}
    A fully injective emergent description is a contradiction in terms under
    Definitions 1 and 2: if $\Pi$ were injective, the effective level would be
    isomorphic to the fundamental level, and there would be no emergence ---
    only a relabelling.

  \section{Framework and Definitions}
  \label{sec:framework}

  Let $\Omega$ be a space of fundamental configurations and $\mathcal{O}$ a
  space of observable states.
  We consider a surjective map $\Pi : \Omega \to \mathcal{O}$.

  \begin{definition}[Observables defined by $\Pi$]
    \label{def:defined-by-Pi}
    The observables of $\mathcal{O}$ are \emph{defined by $\Pi$} if
    $\mathcal{O}$ carries no independent microstate-resolving structure:
    every distinction present in $\mathcal{O}$ is a distinction already present
    in $\Omega$ and transported by $\Pi$, with no additional resolving power
    assigned to $\mathcal{O}$ by an external postulate.
    Equivalently, two configurations $\omega, \omega' \in \Omega$ are
    observationally indiscernible if and only if $\Pi(\omega) = \Pi(\omega')$.
  \end{definition}

  This condition is the minimal formalisation of what it means for observables
  to \emph{arise from} a more fundamental level rather than being postulated
  independently.
  It is weaker than requiring that every structure on $\mathcal{O}$ be
  explicitly constructed from $\Omega$: it only requires that $\mathcal{O}$
  introduces no new resolving power beyond what $\Pi$ already provides.
  Equivalently, Definition~\ref{def:defined-by-Pi} can be rephrased as:
  \emph{$\Pi$ is informationally complete for observables} --- meaning that
  every distinction $\mathcal{O}$ can draw is already drawn by $\Pi$, and
  $\mathcal{O}$ adds no further discriminating capacity of its own.
  Any theory that assigns independent microstate-resolving attributes to
  $\mathcal{O}$ is, by that act, a two-level theory with autonomous observable
  structure, not an emergent one.
  Frameworks where $\mathcal{O}$ carries independent structure --- such as
  standard quantum mechanics or general relativity, which directly postulate
  a Hilbert space or a spacetime manifold --- lie outside this class by
  construction; see Section~\ref{sec:scope} for a precise discussion.

  \begin{example}[Spin model]
    \label{ex:spin}
    Let $\Omega = \{-1,+1\}^N$ be the configuration space of $N$ Ising spins, let
    $\Pi(\sigma) = N^{-1}\sum_i \sigma_i$ be the mean magnetisation, and let
    $\mathcal{O} = \Pi(\Omega) = \{-1, -1+\tfrac{2}{N}, \ldots, 1\}$ be the finite set of attainable
    magnetisation values, so that $\Pi$ is surjective.
    Distinct spin configurations with the same mean magnetisation are
    observationally indiscernible: $\mathcal{O}$ carries no structure that
    distinguishes them.
    Definition~\ref{def:defined-by-Pi} is satisfied.
    Condition~(E3) holds since $|\mathcal{O}| = N+1 \ll 2^N = |\Omega|$ for large $N$.
    Theorem~\ref{thm:main} then guarantees $S_\Pi(\mu) > 0$ for every non-degenerate $\mu$, which is immediate:
    for any non-extremal attainable magnetisation value $m \in (-1,1)$, the fibre $\Pi^{-1}(m)$ contains at
    least two configurations, each of positive mass, and for bulk magnetisation values it contains
    exponentially many configurations.
    The non-injectivity is not merely observed here; it is a structural
    necessity of the emergent description.
  \end{example}

  \begin{definition}[Effectively non-trivial description]
    \label{def:non-trivial}
    The description is \emph{effectively non-trivial} if:
    \begin{itemize}
      \item[(E1)] \textbf{Distinguishability:} $\Pi$ is not constant ---
      there exist $\omega_1, \omega_2 \in \Omega$ such that
      $\Pi(\omega_1) \neq \Pi(\omega_2)$.
      \item[(E3)] \textbf{Structural non-isomorphism:} $\mathcal{O} \not\cong \Omega$
      as structured spaces, where ``structured space'' refers to the full set of
      physically relevant structures induced by $\Pi$ (topology, algebraic
      relations, observables, or dynamical laws).
    \end{itemize}
  \end{definition}

  \begin{remark}
    Condition (E3) is the formal statement that the effective level is
    genuinely distinct from the fundamental level.
    An isomorphism $\mathcal{O} \cong \Omega$ would mean that $\mathcal{O}$ is
    merely a relabelling of $\Omega$, not an emergent structure.
    Note that (E2) --- the statement that $\Pi$ is non-injective --- is
    deliberately absent from Definition~\ref{def:non-trivial}: including it
    would make the main theorem circular.
  \end{remark}

  We now fix the entropy that quantifies the compression of $\Pi$.
  Throughout, $\Omega$ is finite or countable, and a \emph{non-degenerate} measure on $\Omega$ is a probability
  measure $\mu$ with full support: $\mu(\{\omega\}) > 0$ for every $\omega \in \Omega$.
  Let $\nu = \Pi_*\mu$ be the pushforward distribution on $\mathcal{O}$, $\nu(o) = \mu(\Pi^{-1}(o))$.
  Two entropies are naturally associated with $\Pi$, and they must be distinguished.
  The \emph{observable entropy} $H(\nu) = -\sum_o \nu(o)\log\nu(o)$ measures the diversity of the effective
  description; it is the quantity used (normalised by its maximum) in the Bell analysis of~\citep{Beau2026b}.
  The quantity relevant to the present result is the \emph{projection entropy}: the information about $\Omega$
  that the passage to $\mathcal{O}$ erases,
  \begin{equation}
    S_\Pi(\mu) \;\equiv\; H(\mu) - H(\Pi_*\mu) \;=\; \sum_{o \in \mathcal{O}} \nu(o)\, H(\mu_o),
    \label{eq:SPI}
  \end{equation}
  where $\mu_o$ is the conditional distribution of $\mu$ on the fibre $\Pi^{-1}(o)$; the second equality is the
  chain rule for Shannon entropy, so $S_\Pi(\mu)$ is the mean fibre entropy.
  For countable $\Omega$, the conditional-entropy expression is the definition:
  $\sum_o \nu(o)\, H(\mu_o)$ is always well-defined in $[0, \infty]$, and whenever $H(\mu)$ and $H(\Pi_*\mu)$
  are both finite it agrees with the difference $H(\mu) - H(\Pi_*\mu)$ by the chain rule.
  All identities written as entropy differences below are to be read under this finite-entropy condition, or
  equivalently through the conditional form.
  It satisfies $S_\Pi(\mu) \geq 0$, with $S_\Pi(\mu) = 0$ if and only if every fibre of positive measure is a
  singleton, i.e.\ $\Pi$ is $\mu$-almost everywhere injective, and $S_\Pi(\mu) > 0$ if and only if some fibre
  contains more than one point of positive measure.
  For a non-degenerate $\mu$, the last condition is exactly the non-injectivity of $\Pi$.
  The observable entropy carries no such equivalence --- $H(\nu)$ is positive for any non-constant $\Pi$,
  injective or not --- and the conditional form~\eqref{eq:SPI} is the one for which the co-extensivity
  statements of this paper are exact.

  \section{Main Result}
  \label{sec:theorem}

  The proof of Theorem~\ref{thm:main} rests entirely on the following lemma,
  which we state and prove separately to make the logical hinge explicit.

  \begin{lemma}[Injective projection is an isomorphism]
    \label{lem:pivot}
    Let $\Pi : \Omega \to \mathcal{O}$ be surjective and injective, with
    observables defined by $\Pi$ in the sense of Definition~\ref{def:defined-by-Pi}.
    Then $\Pi$ is a structural isomorphism $\mathcal{O} \cong \Omega$.
  \end{lemma}

  \begin{proof}
    Since $\Pi$ is surjective and injective, it is bijective.
    Since all structure on $\mathcal{O}$ is induced from $\Omega$ by $\Pi$
    (Definition~\ref{def:defined-by-Pi}), a bijection that transports all
    structure is a structure-preserving isomorphism.
    Hence $\mathcal{O} \cong \Omega$.
  \end{proof}

  \begin{remark}[Why this lemma is the pivot]
      Lemma~\ref{lem:pivot} is the single logical hinge of the entire paper.
      Its content is deceptively simple: if nothing is lost under $\Pi$ (injectivity)
      and everything in $\mathcal{O}$ comes from $\Omega$ (Definition~\ref{def:defined-by-Pi}),
      then $\mathcal{O}$ and $\Omega$ are the same structured space under different names.
      There is no emergence, only relabelling.
      Condition~(E3) of Definition~\ref{def:non-trivial} forbids precisely this ---
      and thereby forces non-injectivity.
  \end{remark}

  \begin{theorem}[Compression necessity]
    \label{thm:main}
    Let $\Pi : \Omega \to \mathcal{O}$ be surjective, with observables defined by
    $\Pi$ in the sense of Definition~\ref{def:defined-by-Pi} and description
    effectively non-trivial in the sense of Definition~\ref{def:non-trivial}.
    Then $\Pi$ is necessarily non-injective, and for any non-degenerate measure
    $\mu$ on $\Omega$,
    \[
      S_\Pi(\mu) > 0.
    \]
  \end{theorem}

  \begin{proof}
    Suppose for contradiction that $\Pi$ is injective.
    By Lemma~\ref{lem:pivot}, $\mathcal{O} \cong \Omega$ as structured spaces.
    This contradicts condition~(E3).
    Therefore $\Pi$ is not injective.

    Since $\Pi$ is not injective, there exists $o \in \mathcal{O}$ whose fibre $\Pi^{-1}(o)$ contains at least
    two points, each of positive mass because $\mu$ is non-degenerate.
    The conditional distribution $\mu_o$ is then supported on at least two points, so $H(\mu_o) > 0$ and
    $\nu(o) > 0$; hence
    $S_\Pi(\mu) = \sum_{o'} \nu(o')\, H(\mu_{o'}) \geq \nu(o)\, H(\mu_o) > 0$.
  \end{proof}

  \section{Gauge Structure and Co-extensivity}
  \label{sec:gauge}

  \begin{proposition}[Gauge requires non-injectivity]
    \label{prop:gauge}
    In any framework where observables are defined by $\Pi$
    (Definition~\ref{def:defined-by-Pi}), any non-trivial gauge structure
    in $\mathcal{O}$ requires $\Pi$ to be non-injective, and hence $S_\Pi > 0$.
  \end{proposition}

  \begin{proof}
    If $\Pi$ were injective, every fibre $\Pi^{-1}(o)$ would be a singleton.
    A gauge structure arising from $\Pi$ requires a non-trivial action of a structure group $G_\Pi$ on the
    fibres; on singleton fibres only the trivial group acts effectively, so $G_\Pi = \{e\}$ and no non-trivial
    gauge structure can arise from $\Pi$ alone.
    Any non-trivial gauge field would then require an independent postulate at
    the level of $\mathcal{O}$, violating Definition~\ref{def:defined-by-Pi}.
    Non-injectivity, hence $S_\Pi(\mu) > 0$ for every non-degenerate $\mu$, follows.
  \end{proof}

  \begin{remark}[On the converse]
    \label{rem:converse}
    The converse of Proposition~\ref{prop:gauge} does not hold in general:
    non-injectivity does not automatically produce gauge structure.
    For the redundancy encoded in non-trivial fibres to organise into a gauge
    symmetry, additional geometric hypotheses are required: the fibres
    $\Pi^{-1}(o)$ must form a principal bundle over $\mathcal{O}$ with a
    non-trivial structure group $G_\Pi$ admitting a connection.
    This is a result supplementary to Theorem~\ref{thm:main}, not a consequence
    of it.
  \end{remark}

  \begin{corollary}[Co-extensivity]
    \label{cor:co-ext}
    In any projective emergent framework satisfying the hypotheses of
    Theorem~\ref{thm:main},
    \[
      \text{descriptive redundancy}
      \;\Longleftrightarrow\;
      \Pi \text{ non-injective}
      \;\Longleftrightarrow\;
      S_\Pi > 0,
    \]
    and any non-trivial gauge structure requires these properties:
    \[
      \text{non-trivial gauge}
      \;\Longrightarrow\;
      \Pi \text{ non-injective}
      \;\Longleftrightarrow\;
      S_\Pi > 0.
    \]
    Under additional geometric hypotheses on the fibre organisation of $\Pi$
    (principal bundle structure, admissible connection, non-trivial holonomy),
    the redundancy may in turn manifest as a gauge structure.
  \end{corollary}

  This perspective is consistent with the standard interpretation of gauge
  symmetry as a redundancy of description, as formalised in constrained
  Hamiltonian systems~\citep{HenneauxTeitelboim1992} and relational
  approaches to observables~\citep{Rovelli2002}.

  \begin{remark}[Algebraic route to $S_\Pi > 0$ via fibre non-commutativity]
    \label{rem:heisenberg-route}
    Theorem~\ref{thm:main} derives $S_\Pi > 0$ from the structural conditions
    (surjectivity, informational completeness, non-isomorphism) without
    specifying a mechanism for the kernel of~$\Pi$.
    In frameworks where the admissibility structure of the fibre carries a
    Heisenberg-type algebra --- specifically where the fibre generators satisfy
    $[X, Y] = Z \neq 0$ --- a complementary algebraic derivation is available.
    Any projection $\Pi$ with trivial kernel would require the simultaneous
    diagonalisation of $X$ and $Y$, which the non-zero commutator structurally
    forbids; hence $\ker\Pi \neq \{0\}$ and $S_\Pi > 0$ follow directly from
    the algebraic structure of the fibre, independently of the non-isomorphism
    condition (E3).
    The two derivations are complementary: Theorem~\ref{thm:main} is
    framework-independent and answers \emph{what} is structurally forced;
    the algebraic route is mechanism-specific and answers \emph{why} it is
    forced in frameworks whose fibre carries non-commutative structure.
    In such frameworks, structural fluctuations --- the irreducible projective
    residue guaranteed by $\ker\Pi \neq \{0\}$ --- are not contingent
    modelling choices but necessary consequences of fibre non-commutativity,
    providing an independent source for the same conclusion that
    Theorem~\ref{thm:main} reaches by abstract necessity.
  \end{remark}

  \begin{corollary}[Recursive non-injectivity]
    \label{cor:recursive}
    Let $\Omega_0 \xrightarrow{\Pi_1} \Omega_1 \xrightarrow{\Pi_2} \Omega_2$ be a tower
    of surjective projections in which, at each step $i \in \{1,2\}$, the observables of
    $\Omega_i$ are defined by $\Pi_i$ and $\Omega_i \not\cong \Omega_{i-1}$.
    Then each $\Pi_i$ is non-injective, $S_{\Pi_i} > 0$ on non-degenerate measures, and for any non-degenerate
    $\mu$ on $\Omega_0$ the composed projection $\Pi_2 \circ \Pi_1 : \Omega_0 \to \Omega_2$ is non-injective
    with the exact additivity
    \[
      S_{\Pi_2 \circ \Pi_1}(\mu) \;=\; S_{\Pi_1}(\mu) + S_{\Pi_2}\!\left((\Pi_1)_*\mu\right)
      \;>\; S_{\Pi_1}(\mu).
    \]
    The result extends to towers of arbitrary finite depth: the total compression is the sum of the per-stratum
    compressions.
  \end{corollary}

  \begin{proof}
    Non-injectivity of each $\Pi_i$ follows from Theorem~\ref{thm:main} applied at each
    level independently.
    Non-injectivity of the composition is immediate: if $\Pi_1(\omega) = \Pi_1(\omega')$
    for $\omega \neq \omega'$, then $\Pi_2(\Pi_1(\omega)) = \Pi_2(\Pi_1(\omega'))$.
    The entropy identity is the telescoping of Eq.~\eqref{eq:SPI}:
    $H(\mu) - H((\Pi_2 \circ \Pi_1)_*\mu)
    = \left[H(\mu) - H((\Pi_1)_*\mu)\right] + \left[H((\Pi_1)_*\mu) - H((\Pi_2)_*(\Pi_1)_*\mu)\right]$,
    and each bracket is the projection entropy of the corresponding stratum.
    Surjectivity of $\Pi_1$ and full support of $\mu$ make $(\Pi_1)_*\mu$ non-degenerate on $\Omega_1$, so the
    second bracket is strictly positive by Theorem~\ref{thm:main} applied at the second stratum.
  \end{proof}

  \begin{remark}[Self-application of the theorem]
    \label{rem:self-application}
    Corollary~\ref{cor:recursive} expresses a self-application property:
    Theorem~\ref{thm:main} does not merely apply once to the pair $(\Omega, \mathcal{O})$
    posited by a given framework.
    It applies at every pair of adjacent levels in any tower of emergence.
    Non-injectivity therefore accumulates across levels: each emergent description carries its
    own necessary compression, independently of the compressions already realised below it.
    A framework with $k$ genuinely distinct levels of description carries at least $k-1$
    independent non-injective projections, each with strictly positive projection entropy.
  \end{remark}

  \section{Three-Level Structure}
  \label{sec:levels}

  The results of Sections~\ref{sec:theorem} and~\ref{sec:gauge} establish a
  hierarchy with three distinct levels of increasing specificity.

  \textbf{Universal level.}
  Theorem~\ref{thm:main} holds for any surjective $\Pi$ satisfying
  Definitions~\ref{def:defined-by-Pi} and~\ref{def:non-trivial}.
  No dynamical, geometric, or physical hypothesis is required beyond the
  structural conditions on $\Pi$.
  The conclusion --- $\Pi$ non-injective, $S_\Pi > 0$ --- is a structural
  constraint on any framework that takes emergence seriously.

  \textbf{Intermediate level.}
  If the non-trivial fibres of $\Pi$ are organised geometrically --- forming a
  principal $G_\Pi$-bundle with admissible connection --- then the projective
  redundancy can support a gauge structure with holonomy group $G_\Pi$.
  This is an additional geometric hypothesis, not a consequence of
  Theorem~\ref{thm:main} alone.

  \textbf{Concrete realisation.}
  In the pre-geometric projective framework of~\citep{Cosmochrony}, the intermediate
  hypothesis is explicitly realised: the fundamental projection $\Pi$ defines a principal
  bundle with fibre $S^3$, yielding $G_\Pi \supseteq \mathrm{SU}(2) \times \mathrm{U}(1)$.
  The resulting gauge structure underlies the emergence of electroweak
  interactions, while the Born--Infeld saturation bound selects the unique
  admissible infrared dynamics~\citep{Beau2026c}.
  Bell-type correlations arise as the observable signature of $S_\Pi >
  0$~\citep{Beau2026b}.
  The route from non-injectivity to the specific group $\mathrm{SU}(2)$,
  via Born--Infeld parity, Weil representations, and the Hurwitz classification
  of real normed division algebras, is detailed in Appendix~\ref{app:SU2}.

  \section{Towers of Emergence and Structural Confinement}
  \label{sec:towers}

  Corollary~\ref{cor:recursive} elevates the main result from a constraint on a
  single projection to a constraint on any stratified emergent framework.
  Each stratum is subject to Theorem~\ref{thm:main}; the non-injectivities at
  different strata are independent and additive in entropy.
  We make this structure explicit in the framework of~\citep{Cosmochrony,Beau2026a35},
  where the spectral admissibility programme identifies a natural tower, and draw a
  consequence for colour confinement.

  \subsection*{The admissibility tower}

    The spectral admissibility programme of~\citep{Cosmochrony} works with conjugate pairs
    $\{c, q-c\} \subset (\mathbb{Z}/q\mathbb{Z})^*$ as its fundamental observable units.
    Conjugate pairs are colour-neutral by construction: $c + (q-c) = q \equiv 0$, so
    the pair observable $\sigma_{\mathrm{pair}}(n) = \sigma_c(n) \cdot \sigma_{q-c}(n)$
    retains no colour information.
    Leptons, which are colourless, are naturally identified at this level.

    However, each element $c_i$ of a colour triplet
    $\{c_1, c_2, c_3\} \subset (\mathbb{Z}/q\mathbb{Z})^*$~\citep{Beau2026a35}
    already carries a two-dimensional Weil-trajectory space $V_\rho \cong \mathbb{C}^2$ ---
    the spin-$\frac{1}{2}$ structure inherited from its own conjugate pair $\{c_i, q-c_i\}$
    --- together with a colour eigenspace $\tilde{V}_{c_i} \cong \mathbb{C}$ selected
    by Born--Infeld saturation.
    The individual sector $c_i$ therefore carries simultaneously spin-$\frac{1}{2}$ and a
    colour charge: the minimal quantum number content of a quark.

    This identifies a tower of genuinely distinct emergent levels above the pre-geometric
    substrate $\Omega_0$:
    \begin{equation}
      \Omega_0
      \;\xrightarrow{\Pi_1}\;
      \{\text{individual sectors } c_i\}
      \;\xrightarrow{\Pi_2}\;
      \{\text{colour-neutral pairs and triplets}\}
      \;\xrightarrow{\Pi_3}\;
      \{\text{leptons, hadrons}\}.
      \label{eq:tower}
    \end{equation}
    Each arrow satisfies the hypotheses of Theorem~\ref{thm:main}: the target level is
    genuinely distinct from the source (different symmetry group, different admissibility
    structure, different observable content), and its observables are defined by the
    projection.
    By Corollary~\ref{cor:recursive}, each $\Pi_i$ is non-injective with
    $S_{\Pi_i} > 0$.

  \subsection*{Structural confinement}

    The projection $\Pi_2$ in~\eqref{eq:tower} maps coloured sectors $c_i$ to
    colour-neutral pairs and triplets.
    Theorem~\ref{thm:main}, applied to this pair of adjacent levels, forces $\Pi_2$ to
    be non-injective: distinct colour configurations --- distinct assignments of colours to
    the three sectors $c_1, c_2, c_3$ --- map to the same colour-neutral observable.

    This is the structural content of colour confinement: individual coloured sectors are
    not observable because $\Pi_2$ erases colour information.
    Within the projective-emergent setting, confinement is not a dynamical accident
    requiring a separate explanation; it is a structural necessity of the emergence of
    colour-neutral observables from coloured degrees of freedom.

    \begin{remark}[Confinement as projective necessity]
      \label{rem:confinement}
      The present argument does not reproduce the quantitative features of QCD confinement
      (string tension, running coupling, asymptotic freedom).
      It establishes the weaker but more fundamental claim: any framework in which
      colour-neutral hadrons arise as genuinely emergent from coloured sectors --- in the
      sense of Definitions~\ref{def:defined-by-Pi} and~\ref{def:non-trivial} --- must have
      a non-injective colour projection.
      The non-observability of free colour charges is the observable signature of
      $S_{\Pi_2} > 0$ at the colour stratum.
      The quantitative dynamics of confinement is a supplementary question, analogous to
      the question of which specific group $G_\Pi$ organises the fibre redundancy
      (Remark~\ref{rem:converse}).
    \end{remark}

  \subsection*{Recursive self-similarity}

    A structural consequence of Remark~\ref{rem:self-application} is that the compression
    mechanism is self-similar across the tower~\eqref{eq:tower}: the same structural
    argument --- surjectivity, informational completeness, non-isomorphism --- forces
    non-injectivity at each stratum, regardless of which stratum one takes as its own
    ``fundamental level.''
    The non-injectivity of $\Pi_1$ is not a distinguished feature of the deepest level; it
    is the first instance of a pattern that the theorem forces to repeat at every level of
    genuine emergence.
    This self-similarity is not postulated but derived: it is the content of
    Corollary~\ref{cor:recursive} applied to each adjacent pair in the tower.

  \section{Scope and Limits of the Result}
  \label{sec:scope}

  Theorem~\ref{thm:main} is a conditional result.
  Its conclusion holds whenever the hypotheses of
  Definitions~\ref{def:defined-by-Pi} and~\ref{def:non-trivial} are satisfied.
  It is therefore essential to identify the class of theories to which it applies
  and the class to which it does not.

  \textbf{Theories outside the scope.}
  Standard physical frameworks --- quantum mechanics, general relativity,
  standard-model field theories --- are not covered by Theorem~\ref{thm:main}.
  These theories posit their fundamental objects (Hilbert space, spacetime manifold,
  gauge fields) directly at the level $\mathcal{O}$, without deriving them from a
  more primitive space $\Omega$ via a projection.
  The question of non-injectivity does not arise in such frameworks: there is no
  projection $\Pi$ to be injective or otherwise.
  The theorem is silent about them.

  Loop quantum gravity (LQG) deserves explicit discussion, as it occupies an
  intermediate position~\citep{Rovelli2004,Thiemann2007}.
  LQG describes the quantum geometry of the universe --- spin networks and spin
  foams --- as fundamental objects that are themselves part of the observable
  universe, not external to it.
  It does not posit a pre-geometric substrate $\Omega$ from which observables
  are projected; it claims to \emph{be} the fundamental level.
  In this sense LQG is not emergent: it does not fit the projective structure
  of Definition~\ref{def:defined-by-Pi}, and Theorem~\ref{thm:main} does not
  apply to it as a fundamental theory.

  The relevant demarcation criterion is therefore not the presence of a
  singularity but an ontological one:
  \begin{quote}
    \emph{Theorem~\ref{thm:main} applies naturally to frameworks that posit
    a level $\Omega$ structurally external to the observable universe,
      from which observables are projected via $\Pi$.
      A theory that describes the universe at a finer scale --- however fine ---
      without positing such an external level falls outside this class.}
  \end{quote}
  LQG describes the universe at the Planck scale; it does not describe what
  lies outside or before the universe.
  The framework of~\citep{Cosmochrony}, by contrast, posits a pre-geometric substrate that is
  pre-spatial, pre-temporal, and structurally prior to the observable universe:
  it is explicitly in the class to which the theorem applies.

  Note, however, that the semi-classical limit of LQG ---
  the projection $\Pi_\mathrm{sc} : \mathcal{H}_\mathrm{LQG} \to g_{\mu\nu}$
  that recovers classical general relativity --- is itself a non-injective map
  within the class of Definition~\ref{def:defined-by-Pi}, with
  $\Omega = \mathcal{H}_\mathrm{LQG}$ and $\mathcal{O} = \{g_{\mu\nu}\}$.
  Theorem~\ref{thm:main} applies to this projection: general relativity is
  genuinely emergent from LQG, and its projection is necessarily non-injective.
  The two claims --- LQG is not emergent as a fundamental theory, GR is
  emergent from LQG --- are consistent.

  \begin{remark}[Multi-valued maps lie outside the scope]
    \label{rem:multivalued}
    Theorem~\ref{thm:main} assumes a single-valued surjection
    $\Pi : \Omega \to \mathcal{O}$.
    Multi-valued maps $\Phi : \Omega \rightrightarrows \mathcal{O}$,
    in which a single configuration $\omega \in \Omega$ may be associated
    with several distinct observables, fall outside the framework.

    The exclusion is independent of injectivity and follows directly from
    Definition~\ref{def:defined-by-Pi}.
    If $\Phi(\omega)$ contains distinct elements $o_1 \neq o_2$, then
    $\mathcal{O}$ resolves a distinction --- between $o_1$ and $o_2$ ---
    that is not resolved by the fundamental description at $\omega$.
    The observable level then carries a strictly finer discriminating
    structure than the fundamental level, which is incompatible with the
    requirement that $\mathcal{O}$ carry no independent microstate-resolving
    structure.

    Such generative maps describe a structurally different relationship
    between levels: one in which the effective description is more
    expressive than the fundamental one.
    Whether they admit structural necessity results analogous to
    Theorem~\ref{thm:main} is an open question.
  \end{remark}

  \textbf{Theories within the scope.}
  Theorem~\ref{thm:main} applies to any framework where the effective description
  is genuinely derived from a finer-grained level via projection.
  This includes, in particular:
  renormalisation group (RG) approaches, where the projection
  $\Pi_\Lambda : \Omega_\mathrm{UV} \to \mathcal{O}_\mathrm{IR}$ integrates out
  modes above the scale $\Lambda$ and is non-injective by construction;
  causal set models, where macroscopic geometry is an invariant of sets of
  micro-configurations and the mapping is many-to-one;
  spin-foam and spin-network models, where the effective geometrical observables
  arise from averaging over discrete micro-structures;
  and pre-geometric relational frameworks~\citep{Cosmochrony}, where the
  full observable structure is explicitly derived from a pre-geometric substrate
  via a projection $\Pi$.

  In each of these cases, the theorem does not prove non-injectivity from nothing:
  it shows that the non-injectivity already implicit in the framework's own
  definition of ``effective description'' is a logical necessity, not a contingent
  modelling choice.

  \textbf{The key distinction.}
  The result isolates a precise logical boundary:

  \begin{quote}
    \emph{Within the present projective-emergent setting,
      injectivity of $\Pi$ is equivalent to the absence of genuine emergence.}
  \end{quote}

  If $\Pi$ is injective, the effective level is isomorphic to the fundamental
  level: nothing is lost, nothing is emergent.
  If the effective level is genuinely distinct, something is necessarily lost, and
  that loss is precisely $S_\Pi > 0$.

  \section{Categorical and Informational Reformulations}
  \label{sec:cat-info}

  Two reformulations of Theorem~\ref{thm:main} strengthen its connections to
  existing mathematical frameworks.

  \textbf{Categorical form.}
  Fix the pair $(\Omega, \Pi)$.
  Let $\mathbf{Fund}$ be the category with the single object $\Omega$ whose morphisms are the
  \emph{fibre-compatible} maps $f : \Omega \to \Omega$, i.e.\ those satisfying
  $\Pi(\omega) = \Pi(\omega') \Rightarrow \Pi(f(\omega)) = \Pi(f(\omega'))$, and let $\mathbf{Obs}$ be the
  category with the single object $\mathcal{O}$ and all maps $\mathcal{O} \to \mathcal{O}$ as morphisms.
  Every fibre-compatible $f$ descends to a unique map $F_\Pi(f) : \mathcal{O} \to \mathcal{O}$ with
  $F_\Pi(f) \circ \Pi = \Pi \circ f$, and this assignment satisfies
  $F_\Pi(g \circ f) = F_\Pi(g) \circ F_\Pi(f)$ and $F_\Pi(\mathrm{id}_\Omega) = \mathrm{id}_{\mathcal{O}}$:
  $F_\Pi$ is a functor.
  Recall that a functor is \emph{faithful} if it is injective on each hom-set.

  \begin{theorem}[Categorical form of compression necessity]
    \label{thm:cat}
    Under the hypotheses of Theorem~\ref{thm:main}, the functor
    $F_\Pi : \mathbf{Fund} \to \mathbf{Obs}$ is not faithful.
    Conversely, if $F_\Pi$ is faithful then $\Pi$ is injective and, by Lemma~\ref{lem:pivot},
    $\mathcal{O} \cong \Omega$: faithfulness is equivalent to the absence of genuine emergence.
  \end{theorem}

  \begin{proof}
    By Theorem~\ref{thm:main}, $\Pi$ is non-injective: some fibre contains two distinct points
    $\omega_1 \neq \omega_2$.
    Let $g : \Omega \to \Omega$ exchange $\omega_1$ and $\omega_2$ and fix every other point.
    Then $g$ preserves every fibre of $\Pi$, hence is fibre-compatible, and it descends to
    $F_\Pi(g) = \mathrm{id}_{\mathcal{O}} = F_\Pi(\mathrm{id}_\Omega)$ while $g \neq \mathrm{id}_\Omega$.
    Two distinct morphisms of $\mathbf{Fund}$ have the same image, so $F_\Pi$ is not faithful.
    For the converse, suppose $F_\Pi$ is faithful and $\Pi(\omega_1) = \Pi(\omega_2)$ with
    $\omega_1 \neq \omega_2$; the same transposition $g$ would give $F_\Pi(g) = F_\Pi(\mathrm{id}_\Omega)$ with
    $g \neq \mathrm{id}_\Omega$, contradicting faithfulness.
    Hence $\Pi$ is injective, and Lemma~\ref{lem:pivot} gives $\mathcal{O} \cong \Omega$.
  \end{proof}

  \begin{remark}
    Theorem~\ref{thm:cat} identifies genuine emergence with the failure of
    faithfulness in the categorical sense.
    Faithful functors preserve all distinctions of the source; non-faithful
    functors compress them.
    Emergence is precisely this compression, now stated as a categorical necessity.
  \end{remark}

  \textbf{Informational version.}
  Let $p$ be a probability distribution on $\Omega$ and $q = \Pi_* p$ its pushforward on $\mathcal{O}$.
  The information about the fundamental configuration that survives the passage to the effective level is
  $H(q)$; the loss is $H(p) - H(q) = S_\Pi(p) \geq 0$, by the chain rule of Eq.~\eqref{eq:SPI}.
  The data-processing inequality~\citep{CoverThomas2006} guarantees that this loss is non-negative for
  \emph{any} map; it does not force it to be strictly positive.
  Theorem~\ref{thm:main} adds exactly the missing strictness.

  \begin{corollary}[Strict information loss]
    \label{cor:info}
    In any effectively non-trivial emergent framework, every non-degenerate distribution $p$ on $\Omega$
    satisfies
    \[
      S_\Pi(p) \;>\; 0 .
    \]
    If $H(p)$ and $H(\Pi_* p)$ are finite, this is equivalently $H(p) - H(\Pi_* p) = S_\Pi(p) > 0$.
  \end{corollary}

  The distinction with the data-processing inequality should be kept sharp, because it locates the content of
  the theorem.
  The inequality states that no processing can increase information; it is entirely compatible with lossless
  processing ($S_\Pi = 0$, $\Pi$ injective).
  The theorem states that genuine emergence \emph{forces} strictly lossy processing: the hypotheses of
  Definitions~\ref{def:defined-by-Pi} and~\ref{def:non-trivial} exclude the lossless case.
  The claim is therefore not ``a data-processing inequality holds'' but ``genuine emergence forces a strict
  data-processing loss''.

  The informational reading also connects to the renormalisation group.
  The Wilsonian RG projection $\Pi_\Lambda$ is non-injective by construction~\citep{Wilson1975}:
  the same low-energy observable corresponds to infinitely many UV completions.
  Corollary~\ref{cor:info} says that any such effective field theory has
  $S_{\Pi_\Lambda} > 0$: the loss of UV information is
  not a limitation to be overcome but a structural property of the emergent
  description itself.

  This interpretation aligns with the information-theoretic view of
  statistical mechanics and inference, where macroscopic descriptions
  necessarily involve a loss of microscopic detail~\citep{Jaynes1957}.

  \section{Objections and Neighbouring Frameworks}
  \label{sec:objections}

  This section answers, in condensed form, the objections most naturally raised against
  Theorem~\ref{thm:main}, and locates the result relative to neighbouring programmes.
  Each item neutralises one specific objection; none extends the claim.

  \subsection*{``Non-injectivity is obvious for any coarse-graining''}

    This objection is addressed in detail in Section~\ref{sec:discussion}; here we record its external anchor.
    The recent classification of varieties of emergence by Carroll and Parola~\citep{CarrollParola2024} contains
    exactly one variety realised by a one-to-one map, which the authors themselves qualify as a somewhat trivial
    case of emergence; every non-trivial variety in their taxonomy involves a many-to-one map.
    The taxonomy records the pattern as an empirical regularity of the known examples; Theorem~\ref{thm:main}
    converts it into a necessity: within the projective-emergent setting, no non-trivial variety \emph{can} be
    realised by a one-to-one map.

  \subsection*{``The information is still there, merely inaccessible''}

    In coarse-grained statistical mechanics the microstate information remains present in principle --- in the
    microscopic correlations --- and the coarse description is a resource-bounded choice; on this reading the
    loss would be epistemic, not structural.
    Two answers apply.
    First, the loss asserted by Theorem~\ref{thm:main} is a property of the \emph{description} pair
    $(\Omega, \mathcal{O})$, not of any process: nothing happens to $\Omega$, which loses nothing; the strict
    loss $S_\Pi(\mu) > 0$ quantifies what the effective description cannot register.
    Second, Definition~\ref{def:defined-by-Pi} excludes the epistemic reading structurally: if the effective
    level retains an in-principle access to microstate distinctions, then $\mathcal{O}$ carries
    microstate-resolving structure beyond $\Pi$, and the framework lies outside the class considered --- it is a
    two-level description with autonomous observables, not an emergent one.
    The theorem concerns frameworks whose fundamental level is pre-observational: observability itself is
    defined only at the effective level, so the question of recovering the erased information does not arise
    there as a resource problem.

  \subsection*{``To do physics one must recover the Born rule''}

    Correct, and outside the scope.
    The theorem constrains structure (non-injectivity and its entropy), not probability assignments; the
    construction of quantum probability from admissible projections is the object of the wider
    programme~\citep{Cosmochrony, Beau2026b}, and the present result is independent of it.
    The positioning with respect to the ontological-models framework should also be stated.
    Harrigan and Spekkens~\citep{HarriganSpekkens2010} formalise many-to-one relations between ontic states and
    quantum states, and the Pusey--Barrett--Rudolph theorem~\citep{PBR2012} constrains them; generalised
    probabilistic theories~\citep{Barrett2007} posit states and effects in duality as primitives.
    The map considered here is none of these: it relates a pre-quantum configuration space to observable
    outcomes, with the Hilbert-space structure itself emergent, so the PBR constraint --- which operates within
    quantum theory --- is not touched, and the GPT duality is one of the structures whose emergence is in
    question rather than a starting point.

  \subsection*{``A pre-geometric substrate is just non-locality, or superdeterminism''}

    On Bell's taxonomy the classification is accepted, not evaded: a pre-geometric substrate does not satisfy
    Bell's factorisation condition, whose very statement requires spatial separation, so the framework is
    formally outside the Bell-local class --- consistent with experiment~\citep{Beau2026b}.
    The claim is explanatory: the substrate is prior to the geometric context in which locality is defined, and
    the observed violations appear as the signature of the non-injective projection rather than as
    action-at-a-distance within a given geometry.
    Nor is this superdeterminism: no correlation between hidden variables and measurement settings within a
    given ontology is invoked; instead, the subsystem decomposition presupposed by the factorisation condition
    is itself an output of the projection.

  \subsection*{Deterministic substrata and minimal predictive compressions}

    Two neighbouring programmes make the comparison sharp.
    In 't~Hooft's cellular-automaton interpretation~\citep{tHooft2016}, a deterministic substrate underlies
    quantum mechanics via many-to-one equivalence classes of ontological states; if that effective description
    is genuinely emergent in the sense of Definitions~\ref{def:defined-by-Pi} and~\ref{def:non-trivial},
    Theorem~\ref{thm:main} forces the class map to be non-injective --- consistent with, and structurally
    required by, its construction.
    In computational mechanics~\citep{CrutchfieldYoung1989}, the $\varepsilon$-machine is the minimal sufficient
    statistic of a process: the coarsest compression that preserves predictive power, chosen for utility.
    The two results are complementary: computational mechanics quantifies how much can be kept; the present
    theorem states that, under genuine emergence, not everything can.
    Process-theoretic approaches~\citep{Sulis2026} similarly generate spatial and temporal distinctions from a
    pre-spatial level of process actions, sharing the structural stance that the arena of observables is
    generated rather than given.

  \section{Discussion}
  \label{sec:discussion}

  The central result admits a concise reading:
  \begin{quote}
    \emph{Any genuinely emergent physical description necessarily involves
    information compression.
    A fully injective projection would reduce emergence to a mere
    reparametrisation.}
  \end{quote}
  Equivalently, Theorem~\ref{thm:main} can be read as a \emph{no-go theorem
  for faithful emergence}: a surjective projection that is both
  informationally faithful (injective) and genuinely emergent
  ($\mathcal{O} \not\cong \Omega$) cannot exist.
  One of the two conditions must be abandoned; genuine emergence requires
  abandoning injectivity.

  While the connection between emergence and information loss is widely
  recognised in the context of coarse-graining and effective field
  theories, the present result sharpens this intuition into a structural
  necessity: it does not rely on a specific dynamical or statistical
  framework, but follows from minimal assumptions on the relation between
  levels of description.

  The hypotheses are minimal.
  We do not assume a metric, a Hilbert space, a Lagrangian, or any physical
  dynamics.
  We require only that the effective description be defined by projection
  (Definition~\ref{def:defined-by-Pi}) and genuinely distinct from the
  fundamental level (condition~E3).

  The connection to information theory is direct.
  Formally, the projection entropy $S_\Pi(\mu)$ is the conditional entropy of the configuration given the
  observable; what makes it structural rather than epistemic is its origin.
  It is fixed by the fibre structure of $\Pi$ and bounded away from zero by Theorem~\ref{thm:main} for every
  non-degenerate $\mu$: it does not express an observer's limited resources, and no refinement available at the
  effective level can reduce it, because by Definition~\ref{def:defined-by-Pi} the effective level has none.
  Theorem~\ref{thm:main} establishes that this compression is not optional: in
  any genuine emergent framework, $S_\Pi > 0$ is a structural necessity, not a
  contingent feature.

  The result also clarifies the status of gauge symmetry.
  Gauge invariance has long been understood as a redundancy of description.
  Proposition~\ref{prop:gauge} and Remark~\ref{rem:converse} make this precise
  in the projective setting: redundancy is equivalent to non-injectivity, which
  is itself forced by the structure of genuine emergence.
  Gauge symmetry is therefore not a fundamental feature of nature but a
  necessary organisational consequence of any emergent description that is rich
  enough to carry a principal bundle structure.

  \textbf{Projective incompleteness of standard physics.}
  The frameworks $\Lambda$CDM and the Standard Model describe the structure of
  $\mathcal{O}$ with extraordinary precision, but they specify neither the
  fundamental level $\Omega$ nor the projection $\Pi : \Omega \to \mathcal{O}$.
  Yet three independent mechanisms within these frameworks each admit a
  natural reading in terms of a non-injective projection.

  \emph{(i) Cosmological singularity.}
  The metric $g_{\mu\nu}$ ceases to be defined at $t \lesssim t_\mathrm{Planck}$,
  which is structurally consistent with the geometry being an effective
  description arising from a more primitive level.
  If such a level $\Omega$ and projection $\Pi$ are posited and satisfy
  Definitions~\ref{def:defined-by-Pi} and~\ref{def:non-trivial},
  Theorem~\ref{thm:main} forces $\Pi$ to be non-injective.

  \emph{(ii) Renormalisation group.}
  The Standard Model can be naturally interpreted as a Wilsonian effective
  field theory~\citep{Wilson1975} with UV cutoff $\Lambda$.
  Under this interpretation, the map
  $\Pi_\Lambda : \Omega_\mathrm{UV} \to \mathcal{O}_\mathrm{IR}$
  is non-injective: infinitely many UV completions yield the
  same infrared observables.
  Similar compressive structures appear in classical statistical mechanics
  through phase-space coarse-graining, where macroscopic thermodynamic
  variables correspond to vast equivalence classes of microstates.

  \emph{(iii) Quantum measurement.}
  The transition from quantum states to classical outcomes admits a natural
  description as a projection
  $\Pi_\mathrm{meas} : \mathcal{H} \to \mathcal{O}_\mathrm{class}$
  mapping distinct quantum states to the same classical outcome.
  The measurement problem can then be read, in the present language, as the
  question of what determines the structure of $\Pi_\mathrm{meas}$ and why
  it is non-injective.

  These three readings are independent of any hypothesis specific to
  the framework of~\citep{Cosmochrony}.
  They illustrate that the projective structure identified in
  Theorem~\ref{thm:main} is not exotic: it appears wherever an effective
  description is obtained from a finer-grained level.
  What is missing in these frameworks is a structural necessity result:
  while non-injective mappings appear operationally in each case, no general
  theorem establishes that they are unavoidable for emergence itself.
  Theorem~\ref{thm:main} supplies this missing result.
  The following table identifies what is described and what is absent in
  standard physics.

  \begin{center}
    \begin{tabular}{lll}
      \hline
      Level                          & Standard physics & Missing                  \\
      \hline
      Observable space $\mathcal{O}$ & fully described  & ---                      \\
      Fundamental space $\Omega$     & absent           & unknown structure        \\
      Projection $\Pi$               & absent           & law of projection        \\
      Fibres $\Pi^{-1}(o)$           & absent           & microscopic multiplicity \\
      \hline
    \end{tabular}
  \end{center}

  The quantities that standard physics cannot derive --- the fine-structure
  constant $\alpha$, the electron-to-proton mass ratio $m_e/m_p$, the CKM
  mixing angles --- can be naturally interpreted as invariants of $\Pi$ that
  $\mathcal{O}$ alone cannot determine: under this interpretation, they encode
  properties of the fibres $\Pi^{-1}(o)$ that are invisible to any description
  confined to the effective level.

  The incompleteness is not empirical but structural:
  \begin{quote}
    \emph{Standard physics omits the map defining its observables.}
  \end{quote}
  This imposes a precise constraint on any candidate unification or
  quantum-gravity programme: a theory that remains within $\mathcal{O}$ ---
  however sophisticated --- does not complete the description; it extends it.
  Any theory that genuinely completes standard physics must specify $\Omega$,
  $\Pi$, and the non-injective structure of $\Pi$.

  \textbf{Recursive structure and the particle hierarchy.}
  Corollary~\ref{cor:recursive} and Section~\ref{sec:towers} show that the
  structural incompleteness of standard physics is not confined to its deepest level.
  The Standard Model itself contains an internal tower of emergence: leptons and
  hadrons are not fundamental objects but colour-neutral projections of coloured
  degrees of freedom.
  The projection from the quark level to the hadronic level is subject to
  Theorem~\ref{thm:main} in its own right and is therefore necessarily non-injective.
  This is a second instance of structural incompleteness within $\mathcal{O}$ itself:
  the map $\Pi_{\mathrm{SM}} : \{\text{quarks}\} \to \{\text{hadrons}\}$ is a
  non-injective projection whose fibre structure is currently described dynamically
  (QCD) but not derived from a more primitive level.
  Theorem~\ref{thm:main} does not construct this fibre structure; it establishes that
  any framework which derives hadrons as genuinely emergent from quarks must have
  $S_{\Pi_{\mathrm{SM}}} > 0$.
  The quantities that QCD cannot derive from first principles --- the proton-to-electron
  mass ratio $m_p/m_e$, the strong coupling constant $\alpha_s$ at a given scale ---
  can be interpreted, in line with the general argument of this section, as invariants
  of $\Pi_{\mathrm{SM}}$ not accessible from the hadronic level alone.

  \textbf{Singularity elimination does not restore completeness.}
  A natural response is to propose a framework that eliminates the initial
  singularity entirely --- loop quantum cosmology (LQC), pre-Big Bang
  scenarios, bounce cosmologies, or emergent-universe models --- on the
  grounds that if $g_{\mu\nu}$ is everywhere well defined, the cosmological
  argument of Section~\ref{sec:intro} does not apply.
  This response is correct as far as it goes, but it confuses eliminating a
  \emph{signal} of incompleteness with eliminating the incompleteness itself.

  Consider the four main classes.
  LQC replaces the singularity by a quantum bounce at Planck
  density~\citep{Ashtekar2006,Bojowald2001}.
  It has a partial $\Omega$ --- spin-network states --- and a many-to-one
  projection to the effective metric.
  It is more advanced than $\Lambda$CDM in specifying $\Omega$, but $\Pi$
  remains unspecified: the effective dynamics is postulated, not derived from
  the projection structure.
  Pre-Big Bang scenarios (string cosmology~\citep{Gasperini1993},
  ekpyrotic~\citep{Khoury2001}, conformal cyclic~\citep{Penrose2010})
  displace the singularity into a preceding phase, but that phase is itself
  described by an effective geometry --- the question of its own $\Omega$ and
  $\Pi$ is deferred, not answered.
  Bounce cosmologies with everywhere-regular $g_{\mu\nu}$ avoid the
  cosmological argument directly; they remain incomplete through the
  renormalisation and measurement non-injectivities of mechanisms~(ii)
  and~(iii) above.
  Emergent-universe models (Hartle--Hawking~\citep{HartleHawking1983},
  tunnelling from nothing~\citep{Vilenkin1982}) replace
  the singularity by a quantum-to-classical transition --- which is itself a
  non-injective projection from $\mathcal{H}$ to a classical geometry.

  In every case, the structure of the argument is the same:
  \begin{quote}
    \emph{Eliminating the singularity removes one signal of projective
    incompleteness.
    It does not supply $\Omega$, $\Pi$, or the non-injective structure of $\Pi$.
    A theory without a singularity but without a fundamental level is simply
    a theory whose limit of validity has not yet been reached within the
    observed range.}
  \end{quote}
  Completeness in the projective sense requires specifying what lies above
  $\mathcal{O}$, not merely extending $\mathcal{O}$ to a regime where it no
  longer breaks down.

  The three reformulations --- structural (Theorem~\ref{thm:main}), categorical
  (Theorem~\ref{thm:cat}), and informational
  (Corollary~\ref{cor:info}) --- are equivalent in content but address different
  audiences and connect to different existing bodies of work.
  Their convergence on the same conclusion strengthens the claim that
  non-injectivity is a universal structural feature of emergent descriptions,
  not an artefact of any particular formalism.

  \textbf{On the non-triviality of the result.}
  A natural objection is that non-injectivity is ``obvious'' in any
  coarse-graining procedure.
  This objection conflates two distinct claims.
  It is indeed obvious that a specific coarse-graining --- such as the spin
  model of Example~\ref{ex:spin} --- produces a non-injective map.
  What Theorem~\ref{thm:main} establishes is different: it shows that
  \emph{any} framework satisfying Definitions~\ref{def:defined-by-Pi}
  and~\ref{def:non-trivial} must be non-injective, independently of the
  specific form of $\Pi$.
  The theorem is a constraint on the class of theories, not an observation
  about a particular model.
  Its content is the logical equivalence
  $\text{genuine emergence} \Leftrightarrow \text{non-injectivity}$
  within the projective-emergent setting, which rules out the possibility of
  a faithful emergent description --- something that is not guaranteed by
  pointing to examples of coarse-graining.
  The present result does not introduce a new model of emergence, but
  establishes a structural constraint that any such model must satisfy:
  within the present projective-emergent framework, emergence without
  non-injectivity is impossible.

  Within the present projective-emergent setting, the proved implication is
  \begin{equation}
    \text{genuine emergence}
    \;\Longrightarrow\;
    \text{information loss}
    \;\Longleftrightarrow\;
    \Pi \text{ non-injective}
    \;\Longleftrightarrow\;
    S_\Pi > 0 .
    \label{eq:equivalence}
  \end{equation}
  The converse requires the effective level to be taken as a genuine quotient description: $\mathcal{O}$ must
  carry no microstate-resolving structure beyond $\Pi$ (Definition~\ref{def:defined-by-Pi}), and it must not be
  structurally isomorphic to $\Omega$ (condition~E3).
  Under these hypotheses, a non-injective projection is precisely the structural mark of a genuinely emergent
  level.
  Without them, non-injectivity alone is only compression, not yet emergence.
  A fully injective effective theory, by contrast, is not an emergent theory but a renaming of the fundamental
  level.

  \subsection*{Interpretive outlook}

  The following reading goes beyond what the theorem proves and is offered as interpretation, with its status
  explicitly labelled.
  If genuine emergence structurally forces information loss, then the effective theories of standard physics
  --- $\Lambda$CDM and the Standard Model --- are naturally read not as approximations of a more exact theory
  of the same observables, but as quotient descriptions whose free constants are invariants of an unspecified
  projection (Section~\ref{sec:discussion}).
  On this reading, the persistent resistance of these constants to derivation from within the effective level
  is not an accident of theoretical progress but a structural feature: by
  Definition~\ref{def:defined-by-Pi}, the effective level has no resources to resolve its own fibres.
  This suggests a reorientation of the reduction debate: the question ``can the effective theory be derived
  from a fundamental one?'' is well posed only together with the projection that connects them, and the
  theorem shows that any such projection worth the name must forget.
  This interpretation is speculative-structural: it is consistent with the theorem and motivated by it, but
  the identification of a specific $\Omega$ and $\Pi$ for standard physics is an open programme, not a result
  of this paper.

  \appendix

  \section{On the Route from Non-Injectivity to the Standard Model Gauge Group}
  \label{app:SU2}

  Theorem~\ref{thm:main} guarantees the existence of non-trivial fibres but
  says nothing about their form.
  The identification of the structure group $G_\Pi$ with the Standard Model gauge
  group $G_{\mathrm{SM}} = \mathrm{SU}(3) \times \mathrm{SU}(2) \times \mathrm{U}(1)$
  is a separate result, established in~\citep{Cosmochrony} and subsequent
  papers via a chain of constraints absent from the universal theorem.
  We sketch the main steps for $\mathrm{SU}(2)$, then indicate how the same
  mechanism extends to the full gauge group.

  \subsection*{The SU(2) identification}

    The route to $\mathrm{SU}(2)$ proceeds in four steps, now entirely analytical.
    First, the Born--Infeld parity constraint~\citep{Beau2026a22} fixes the
    minimal fibre structure to the involution $c \leftrightarrow q - c$ on the
    Weil representation space of $\mathrm{Heis}_3(\mathbb{Z}/q\mathbb{Z})$.
    Second, the requirement that the representation algebra be associative,
    combined with the Hurwitz classification of real normed division algebras
    ($\mathbb{R}$, $\mathbb{C}$, $\mathbb{H}$, $\mathbb{O}$), selects
    $\mathbb{H}$ as the minimal associative non-commutative structure, with
    $\dim_{\mathbb{R}} \mathrm{Im}\,\mathbb{H} = 3$ admissible directions~\citep{Beau2026a26}.
    Third, the quaternionic rigidity theorem~\citep{Beau2026a31} proves that
    every admissible morphism on $\mathrm{Heis}_3(\mathbb{Z}/q\mathbb{Z})$ necessarily
    factors through $\mathfrak{su}(2)$, and $r_{\mathrm{eff}} = 3$ identifies the spin-$1$ carrier
    $H_{\mathrm{eff}} \cong \mathrm{Sym}^2(V_\rho)$, with $d_\rho = 2$ the spin-$\tfrac{1}{2}$ square-root
    doublet~\citep{Beau2026a33}.
    The exclusion of $\mathrm{SO}(3)$ is structural: quotienting the central
    $\mathbb{Z}_2$ of $\mathrm{SU}(2)$ destroys the $4\pi$-periodicity required
    for spinorial holonomy.
    Fourth, the Born--Infeld parity anti-linearity forces an equatorial constraint
    on the admissible fibre, from which the invariant covariance ratio
    $\lambda_1/\lambda_2 = 2$ is derived analytically as the squared
    symmetric-tensor normalisation of the Carnot-degree-2 central generator
    $Z = [X, Y]$~\citep{Beau2026a34}.

    These four steps close the following analytical chain:
    \begin{equation}
      \text{emergence} \;\Rightarrow\; \mathfrak{su}(2)\;\text{\citep{Beau2026a31}}
      \;\Rightarrow\; d_\rho = 2\;\text{\citep{Beau2026a33}}
      \;\Rightarrow\; \lambda_1/\lambda_2 = 2\;\text{\citep{Beau2026a34}, proved}.
      \label{eq:SU2chain}
    \end{equation}
    Each implication in~\eqref{eq:SU2chain} is a proved theorem.

  \subsection*{Extension to the full gauge group}

    The $\mathrm{U}(1)$ factor corresponds to global phase rotations within the
    complex fibre $\mathbb{C}^q$, which lie in the kernel of $\Pi$ by the phase
    invariance of $\sigma_c(n)$~\citep{Beau2026a31}.
    Its identification is unconditional.

    The $\mathrm{SU}(3)$ factor arises from a structurally distinct mechanism: the colour-triplet
    co-admissibility of $\{c_1, c_2, c_3\} \subset (\mathbb{Z}/q\mathbb{Z})^*$ satisfying
    $c_1 + c_2 + c_3 \equiv 0 \pmod{q}$ and pairwise non-conjugacy~\citep{Beau2026a35}.
    On the colour-adapted Cayley graph, the triplet construction closes analytically: the group of
    Born--Infeld-compatible fibre-preserving transformations of the combined effective space
    $\tilde{V}_{\mathrm{color}} \cong \mathbb{C}^3$ is exactly $\mathrm{SU}(3)$, with the alternatives
    ($\mathrm{U}(3)$, $\mathrm{SO}(3)$, $\mathrm{U}(2)$, $G_2$) excluded~\citep{Beau2026a35}.
    On the standard graph, the identification is supported at the rank, averaged, effective-exponent, and
    numerical levels~\citep{Beau2026a35, Beau2026a36}; the pointwise finite-$q$ profile equality remains the
    condition that promotes the centre-neutral triplet structure to the full $\mathrm{SU}(3)$ colour
    interpretation.
    The composite admissibility fixed point is therefore
    $G_{\mathrm{SM}} = \mathrm{SU}(3) \times \mathrm{SU}(2) \times \mathrm{U}(1)$,
    with the $\mathrm{U}(1)$ and $\mathrm{SU}(2)$ factors closed analytically, the colour-adapted
    $\mathrm{SU}(3)$ sector closed in its adapted setting, and the standard-graph $\mathrm{SU}(3)$ reading
    conditional on the remaining pointwise co-admissibility hypothesis.

  \subsection*{Complementarity with Theorem~\ref{thm:main}}

    None of the steps above --- Born--Infeld dynamics, Weil representations,
    Hurwitz theorem, quaternionic rigidity, colour triplet co-admissibility ---
    appears in Theorem~\ref{thm:main}.
    The universal theorem supplies the necessary condition (non-trivial fibres,
    $S_\Pi > 0$); the framework-specific constraints supply the sufficient
    conditions that fix the fibre geometry to $G_{\mathrm{SM}}$.
    This selection is model-dependent: other emergent frameworks may yield
    different structure groups.
    The two results are complementary and non-redundant.

  \section*{Statements and declarations}

  % Acknowledgements, funding and competing-interest statements are on the separate title page,
  % per the journal's double-blind procedure.
  Large language models were used as a supportive tool for refining language, structure, and internal
  consistency during the development of this manuscript; all conceptual contributions, theoretical choices,
  and interpretations remain the sole responsibility of the author(s).

  \bibliographystyle{apalike}
  \bibliography{cosmochrony-bibliography, external-refs}

\end{document}
